---
title: "Building a ggplot!"
author: "Jack Talboys"
date: "20/08/2020"
output: html_document
---
We're going to use data from #TidyTuesday, for anyone not aware check out it's [github page](https://github.com/rfordatascience/tidytuesday) for more information. Essentially a weekly dataset is released, and anyone can create a visualization however they want, and show it to the world using the #TidyTuesday on twitter (a really good way to get your data visualizations out there!).

We're going to look at the data on beer from week 14!

Start by pulling the data using the `{tidytuesdayR}` package then assigning each individual data frame to its own object.

```{r}
# install.packages("tidytuesdayR")
library(tidytuesdayR)
beer <- tt_load(2020, week = 14)

beer_states <- beer$beer_states
beer_taxed <- beer$beer_taxed
brewer_size <- beer$brewer_size
brewing_materials <- beer$brewing_materials
```

I've only got half an hour so just going to look at improving a simple visualization that doesn't require any pre-processing of the data.
The figures come from the `beer_states` dataset:

```{r}
head(beer_states)
```

Starting with a simple plot of the number of barrels produced by year. This is where we finally get to use `{ggplot2}` - there's some steps that will be necessary throughout all of your ggplot's so I'll introduce those first.

```{r}
# first load (and maybe install if you need to) the ggplot2 package
#install.packages("ggplot2")
library(ggplot2)

# All plots begin with a call to ggplot
ggplot(data = beer_states)

```

This returns a blank plot, and it's from here that we start to build our plot.

We decided to plot number of barrels produced by year, to specify this we can add to our `ggplot` call by passing information through to the `mappings` argument via the `aes` function as follows:

```{r}
ggplot(data = beer_states,
       mapping = aes(x = year,
                     y = barrels))
```

So this time we get a blank plot, but our axis are in place with the correct range of values and labels taken straight from the column names.


# TODO:
* CONVERT YEAR TO FACTOR
* CHANGE BARRELS FROM SCIENTIFIC NOTATION
* CHOOSING A GEOM
* CHANGING COLOURS/FILLS